\mode*

More concretely, after completing the course, you should be able to:
\begin{frame}
\begin{itemize}
  \item \emph{evaluate} the usability of security solutions and \emph{suggest} 
    improvements that improve usability and security.
  \item \emph{evaluate} threats, possible protection mechanisms and 
    \emph{design} a high-level approach to protection which considers 
    usability.
  \item \emph{navigate} the field of information security, \emph{distinguish} 
    your own limits and where to search for solutions, \eg experts or published 
    research results that are relevant to the solution of a given problem.
  \item \emph{analyse and apply} the results of published research in the 
    security field.
  \item \emph{apply} the Swedish Civil Contingency Agency's Framework for 
    Information Security Management Systems (ISO 27000) to \emph{analyse, assess 
      and improve} the information security in an organization.
\end{itemize}
\end{frame}
The course has a variety of learning sessions designed to ensure that you learn 
these \acp{ILO}.
Each such session has a set of further specified \acp{ILO} that will help you 
achieve the \acp{ILO} above.

The grades will be based on the following grading criteria.
\begin{frame}
\begin{description}
  \item[Grade E] You fulfil all the \acp{ILO} above.
    You should have identified a relevant problem, and given a solution to it.
    It must be a viable solution, however gaps and mistakes are allowed, if 
    they don't render your solution unusable.

  \item[Grade C] You fulfil the criteria for E.
    Additionally, your evaluations and designs are \emph{good} with \emph{some 
      base} in theory and, where applicable, the research literature.
    Gaps and errors are allowed if they only render your solution less optimal.

  \item[Grade A] You fulfil the criteria for C.
    However, your evaluations and designs must be \emph{extensive} and 
    \emph{well-founded} in theory and, where applicable, the research 
    literature.
    Gaps and errors are not allowed in the solution unless they have been 
    properly addressed and you have given a suggestion on an approach to how to 
    start resolve the issue.
\end{description}
\end{frame}
The grades B, C and D are intermediary grades.

